Emergence of the Southeast Asian islands as a driver for Neogene cooling

### 2020-08-12 Manuscript/Supplementary_Information.tex
revisions checkpoint

A related issue is that our analysis focuses on the weathering of Ca and Mg silicates. Sources of alkalinity to the ocean include the four major cations of Ca, Mg, Na and K. A common approach for the long-term carbon cycle follows that of \citet{Berner1984a} where an emphasis is placed on Ca and Mg with the weathering of Na and K silicates considered to not be a long-term sink of \COtwo. The reasoning behind this assumption is that while weathering of these silicates can lead to transient uptake of \COtwo, this uptake is balanced by release. The ultimate mechanism through which silicate weathering sequesters carbon is the incorporation of Ca+Mg carbonates into the lithosphere. While alkalinity generated by the weathering of silicate Na and K consumes \COtwo, that alkalinity is removed through processes such as reverse weathering reactions and exchange for Ca that ultimately precipitates as carbonate \cite{Sun2016a}. An alternative approach for estimating carbon sequestration associated with silicate weathering that sought to incorporate all four cations was taken by \citet{France-Lanord1997a} who used a formulation of $\Delta CO_2 = \Delta Mg + \Delta Ca + 0.15 \Delta Na + 0.1 \Delta K$. This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^2+$ and Mg$^2+$ eventually forms carbonates and applies an estimate that 20\% of K$^+$ and 30\% of Na$^+$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance). Overall, uncertainty in the sinks for K and Na and the processes that control them make them difficult to parameterize and model. Given that we are modeling steady-state solutions and the ultimate sequestration of carbon through carbonate formation, we make the simplification of restricting the modeling of carbon sequestration to that associated with Ca and Mg delivery. Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the overall results would be similar.

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{Figures/lithology_composition.pdf}
    \caption{Ca, Mg, Na, and K concentrations within mafic, intermediate, and felsic lithologies, as calculated based on the mean of data compiled from EarthChem.}
    \label{fig:lithology_composition}
\end{figure}

A related issue is that our analysis focuses on the weathering of Ca and Mg silicates. Sources of alkalinity to the ocean include the four major cations of Ca, Mg, Na and K. A common approach for the long-term carbon cycle follows that of \citet{Berner1984a} where an emphasis is placed on Ca and Mg with the weathering of Na and K silicates considered to not be a long-term sink of \COtwo. The reasoning behind this assumption is that while weathering of these silicates can lead to transient uptake of \COtwo, this uptake is balanced by release. The ultimate mechanism through which silicate weathering sequesters carbon is the incorporation of Ca+Mg carbonates into the lithosphere. While alkalinity generated by the weathering of silicate Na and K consumes \COtwo, that alkalinity is removed through processes such as reverse weathering reactions and exchange for Ca that ultimately precipitates as carbonate \cite{Sun2016a}. An alternative approach for estimating carbon sequestration associated with silicate weathering that sought to incorporate all four cations was taken by \citet{France-Lanord1997a} who used a formulation of $\Delta CO_2 = \Delta Mg + \Delta Ca + 0.15 \Delta Na + 0.1 \Delta K$. This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^2+$ and Mg$^2+$ eventually forms carbonates and applies an estimate that 20\% of K$^+$ and 30\% of Na$^+$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance). Overall, uncertainty in the sinks for K and Na and the processes that control them make them difficult to parameterize and model. Given that we are modeling steady-state solutions and the ultimate sequestration of carbon through carbonate formation, we make the simplification of restricting the modeling of carbon sequestration to that associated with Ca and Mg delivery. Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but given that mafic lithologies have significantly higher Ca+Mg concentrations than intermediate/felsic lithologies relative to the slightly higher Na+K concentration of intermediate/felsic lithologies compared to mafic lithologies (Fig. \ref{fig:lithology_composition}), the importance of the emergence of the SEAI on Neogene \pCOtwo would remain unchanged.
